---
title: "DATA 607 Final Project"
author: "Susanna Wong"
date: "2023-04-30"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Data {.tabset}

Crime Data: Use publicly available crime data from the New York City Open Data.
The data includes information such as the type of crime, the location of the crime, the date of the crime, information of perpetrators.

NY Borough Population Size: Webscrap population size of each NY borough from CityPopulation.
The website contains the population of each NY borough according to census results.
<https://www.citypopulation.de/en/usa/newyorkcity/>

## NYC Open Data

| Column Name            | Description                                                                                                                                                          | Data Type |
|-------------------|-------------------------------|----------------------|
| ARREST_KEY             | Randomly generated persistent ID for each arrest                                                                                                                     | int       |
| ARREST_DATE            | Exact date of arrest for the reported event                                                                                                                          | chr       |
| PD_CD                  | Three digit internal classification code (more granular than Key Code)                                                                                               | int       |
| PD_DESC                | Description of internal classification corresponding with PD code (more granular than Offense Description)                                                           | chr       |
| KY_CD                  | Three digit internal classification code (more general category than PD code)                                                                                        | int       |
| OFNS_DESC              | Description of internal classification corresponding with KY code (more general category than PD description)                                                        | chr       |
| LAW_CODE               | Law code charges corresponding to the NYS Penal Law, VTL and other various local laws                                                                                | chr       |
| LAW_CAT_CD             | Level of offense: felony, misdemeanor, violation                                                                                                                     | chr       |
| ARREST_BORO            | Borough of arrest. B(Bronx), S(Staten Island), K(Brooklyn), M(Manhattan), Q(Queens)                                                                                  | chr       |
| ARREST_PRECINCT        | Precinct where the arrest occurred                                                                                                                                   | int       |
| JURISDICTION_CODE      | Jurisdiction responsible for arrest. Jurisdiction codes 0(Patrol), 1(Transit) and 2(Housing) represent NYPD whilst codes 3 and more represent non NYPD jurisdictions | int       |
| AGE_GROUP              | Perpetrator's age within a category                                                                                                                                  | chr       |
| PERP_SEX               | Perpetrator's sex description                                                                                                                                        | chr       |
| PERP_RACE              | Perpetrator's race description                                                                                                                                       | chr       |
| X_COORD_CD             | Midblock X-coordinate for New York State Plane Coordinate System, Long Island Zone, NAD 83, units feet (FIPS 3104)                                                   | int       |
| Y_COORD_CD             | Midblock Y-coordinate for New York State Plane Coordinate System, Long Island Zone, NAD 83, units feet (FIPS 3104)                                                   | int       |
| Latitude               | Latitude coordinate for Global Coordinate System, WGS 1984, decimal degrees (EPSG 4326)                                                                              | num       |
| Longitude              | Longitude coordinate for Global Coordinate System, WGS 1984, decimal degrees (EPSG 4326)                                                                             | num       |
| Lon_Lat                | Georeferenced Point Column based on Longitude and Latitude fields                                                                                                    | chr       |
| Zip.Codes              | Zip code of arrest                                                                                                                                                   | int       |
| Community.Districts    |                                                                                                                                                                      | int       |
| Borough.Boundaries     |                                                                                                                                                                      | int       |
| City.Council.Districts |                                                                                                                                                                      | int       |
| Police.Precincts       |                                                                                                                                                                      | int       |

```{r}
data_2022 <- read.csv("https://raw.githubusercontent.com/suswong/DATA-607-Final-Project/main/NYPD_Arrests_Data__Historic_-2022.csv", header = T, na.strings = "NA")

data_2021 <- read.csv("https://raw.githubusercontent.com/suswong/DATA-607-Final-Project/main/NYPD_Arrests_Data__Historic_-2021.csv", header = T, na.strings = "NA")

data_2020 <- read.csv("https://raw.githubusercontent.com/suswong/DATA-607-Final-Project/main/NYPD_Arrests_Data__Historic_-2020.csv", header = T, na.strings = "NA")

data_2019 <- read.csv("https://raw.githubusercontent.com/suswong/DATA-607-Final-Project/main/NYPD_Arrests_Data__Historic_-2019.csv", header = T, na.strings = "NA")

data_2018 <- read.csv("https://raw.githubusercontent.com/suswong/DATA-607-Final-Project/main/NYPD_Arrests_Data__Historic_2018.csv", header = T, na.strings = "NA")

data_2018_2022 <- rbind(data_2018,data_2019,data_2020,data_2021,data_2022)
```

## NY Population

```{r}
library(rvest)
web = read_html("https://www.citypopulation.de/en/usa/newyorkcity/")
HTML_df <- html_table(web) 
population <- HTML_df[[1]]

library(DT)
datatable(population)
```

# Tidying Data {.tabset}

## NYC Arrest Data {.tabset}

Below are some of things that needs to be tidied for the analysis:

-   Remove columns that are not needed in the analysis

-   Rename values in the 'ARREST_BORO' column

-   Change format of 'ARREST_DATE' to date format

-   Missing values

### Remove columns

Use `str()` to check the structure of each dataset.

We do not need the following columns in our analysis:

-   ARREST_KEY

-   PD_CD

-   KY_CD

-   LAW_CODE

-   ARREST_PRECINCT

-   JURISDICTION_CODE

-   Community.Districts

-   Borough.Boundaries

-   City.Council.Districts

-   Police.Precincts

```{r}
str(data_2018_2022)

library(dplyr)
data_2018_2022 <- data_2018_2022 %>%
  select(-ARREST_KEY,-KY_CD, -PD_CD, -LAW_CODE, -ARREST_PRECINCT, -JURISDICTION_CODE, -Community.Districts, -Borough.Boundaries, -City.Council.Districts, -Police.Precincts)

```

### Rename Values in 'ARREST_BORO' and 'LAW_CAT_CD'

The 'ARREST_BORO' column contains the borough in which the arrest was made.
The 'LAW_CAT_CD' column contains the level of offense: felony, misdemeanor, violation.
However, those values are in abbreviation.

```{r}
data_2018_2022$ARREST_BORO[grep("B", data_2018_2022$ARREST_BORO)] <- "Bronx"
data_2018_2022$ARREST_BORO[grep("K", data_2018_2022$ARREST_BORO)] <- "Brooklyn"
data_2018_2022$ARREST_BORO[grep("M", data_2018_2022$ARREST_BORO)] <- "Manhattan"
data_2018_2022$ARREST_BORO[grep("Q", data_2018_2022$ARREST_BORO)] <- "Queens"
data_2018_2022$ARREST_BORO[grep("S", data_2018_2022$ARREST_BORO)] <- "Staten Island"

data_2018_2022$LAW_CAT_CD[grep("F", data_2018_2022$LAW_CAT_CD)] <- "Felony"
data_2018_2022$LAW_CAT_CD[grep("M", data_2018_2022$LAW_CAT_CD)] <- "Misdemeanor"
data_2018_2022$LAW_CAT_CD[grep("V", data_2018_2022$LAW_CAT_CD)] <- "Violation"
```

### Change format of 'ARREST_DATE'

The 'ARREST_DATE' needs to be converted to date format.

```{r}
library(lubridate)
data_2018_2022$ARREST_DATE <- as.Date(data_2018_2022$ARREST_DATE, format = "%m/%d/%Y" )

data_2018_2022$YEAR <- as.numeric(format(data_2018_2022$ARREST_DATE,'%Y'))
data_2018_2022$MONTH <- as.numeric(format(data_2018_2022$ARREST_DATE,'%m'))

data_2018_2022$WKDAY <-  wday(data_2018_2022$ARREST_DATE, label=TRUE)
```

### Missing Values

Use `colSums(is.na())` to check for missing values.
There are 2093 missing values in the 'Zip.Codes' column.

```{r}
colSums(is.na(data_2018_2022))

data_2018_2022 <- data_2018_2022[complete.cases(data_2018_2022),]

data_2018_2022 <- data_2018_2022 %>%
  filter(OFNS_DESC !="") %>%
  filter(OFNS_DESC != "(null)")

colSums(is.na(data_2018_2022))
```

### Duplicates

Check for duplicates

```{r}
distinct(data_2018_2022)
```

## NY Population by Borough {.tabset}

To calculate the crime rate by borough, we need the population of each borough.
We will only need the population census in 2020.

Below are some of things that needs to be tidied for the analysis:

-   Remove columns that are not needed in the analysis

-   Rename column labels

-   Change format of the population column to numeric

### Remove columns

Removes the other years' population size except for 2020

```{r}
population_boro <- population[, -c(2:5,7:8)] #Removes the other years' population size except for 2020

population_boro <- as.data.frame(population_boro)
datatable(population_boro)
```

### Rename Column and Values

```{r}
colnames(population_boro) <- c("Borough", "Population_Census_2020")
population_boro$Borough[grep("Bronx", population_boro$Borough)] <- "Bronx"
population_boro$Borough[grep("Brooklyn", population_boro$Borough)] <- "Brooklyn"
population_boro$Borough[grep("Staten Island", population_boro$Borough)] <- "Staten Island"
population_boro$Borough[grep("Manhattan", population_boro$Borough)] <- "Manhattan"

str(population_boro)

population_boro$Population_Census_2020 <- as.numeric(gsub(",","",population_boro$Population_Census_2020))

```

# Interactive Map Visusalization

```{r}
library(dplyr)
library(leaflet)
library(RColorBrewer)
arrests<-rename(data_2022,lat=Latitude,lng=Longitude, radius = 5)
info<-paste("<b>Offense: </b>",arrests$PD_DESC,"<br>",
            "<b>Perpetrator Gender: </b>",arrests$PERP_SEX,"<br>",
            "<b>Perpetrator Age: </b>",arrests$AGE_GROUP,"<br>",
            "<b>Perpetrator Race: </b>",arrests$PERP_RACE,sep = "")

pal<-colorFactor(c("blue","darkgreen","red"),arrests$LAW_CAT_CD)

arrests_map <- arrests %>%
  leaflet()%>%
  addTiles()%>%
    addCircleMarkers(clusterOptions=markerClusterOptions(),
               popup=info,color=pal(arrests$LAW_CAT_CD)) 
# %>%
#     addLegend(labels=c("Free - Unlimited","Free - Limited","Partner"),colors=c("blue","darkgreen","red"))
arrests_map
```

# Daily Crimes in New York from 2018 – 2022

Overall, there is a decreaase in the number of crimes from 2018 to mid 2020 and a increase in the number of crimes from mid 2020 to end of 2022. 
```{r}
crime_daily <- data_2018_2022 %>%
  group_by(ARREST_DATE) %>%
  summarize(count = n()) %>%
  arrange(ARREST_DATE)

library(ggplot2)
library(scales)
ggplot(crime_daily, aes(x = ARREST_DATE, y = count)) +
  geom_line(color = "#F2CA27", size = 0.1) +
  geom_smooth(color = "#1A1A1A") +
  # fte_theme() +
  scale_x_date(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  labs(x = "Date of Crime", y = "Number of Crimes", title = "Daily Crimes in New York from 2018 – 2022")

```

# Top 10 Offenses in NY {.tabset}

## Datatable

Out of all the offenses, assault 3 & related offenses is the highest. 
```{r}
library(DT)
offense <- data_2018_2022 %>%
  group_by(OFNS_DESC) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

datatable(offense)
```

## Visualization

```{r}
library(ggplot2)

offense %>% 
    arrange(desc(count)) %>%
    slice(1:10) %>%
    ggplot(., aes(x=reorder(OFNS_DESC, count), y=count))+
              geom_bar(stat='identity', fill="blue") + coord_flip() + xlab("Offense")+ ylab("Count") +  
  ggtitle("Top 10 Offense in NY")
```

# Crime Rate by Borough and Year {.tabset}

Calculate the crime rate for each borough by dividing the number of arrests in each borough by its population.
This will allow us to compare crime rates on an equal basis across different boroughs.

## Datatable

```{r}
# Count the number of arrests in each borough by year
borough_by_year <- data_2018_2022 %>%
  group_by(YEAR, ARREST_BORO) %>%
  summarise(count = n()) 

# Create a new column for the population size for each borough
borough_by_year$population <- 
  with(borough_by_year, ifelse(ARREST_BORO == "Brooklyn", population_boro[2,2],
                         ifelse(ARREST_BORO == "Bronx", population_boro[1,2],
                        ifelse(ARREST_BORO == "Queens", population_boro[4,2],
                         ifelse(ARREST_BORO == "Manhattan", population_boro[3,2],population_boro[5,2]      )))))

# Rename column names
colnames(borough_by_year) <- c("Year" ,"Borough", "count", "population")

# Calculate the crime rate per year and borough
borough_by_year <- borough_by_year %>%
  mutate(crime_rate = (count/population)*1000)

datatable(borough_by_year,options = list(pageLength = 5, dom = 'tip'))
```

## Visualization

```{r}
ggplot(borough_by_year, aes(fill= Borough, y= crime_rate, x=Year)) + 
  ggtitle("Crime Rate Per 1000 People") + theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
    geom_bar(position="dodge", stat="identity") 
# + scale_y_continuous(labels = scales::percent) 
```

```{r}
ggplot(borough_by_year, aes(fill= reorder(Borough, crime_rate), y= crime_rate, x=Year)) + 
  ggtitle("Crime Rate Per 1000 People") + theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
    geom_bar(position="dodge", stat="identity") 
# + scale_y_continuous(labels = scales::percent) 

```

```{r}
ggplot(borough_by_year, aes (x = factor(crime_rate), y = crime_rate, fill = Borough))+
    geom_bar(stat="identity", na.rm = TRUE)+
    facet_wrap(~Year, ncol =3, scales = "free_x", drop = TRUE) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )
```

```{r}
ggplot(borough_by_year, aes(x = Year, y = crime_rate, colour = Borough)) +
  geom_smooth() + scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) 
```



## Top Offense Per Borough
```{r}
crime_by_borough <- data_2018_2022 %>%
  group_by(OFNS_DESC,ARREST_BORO) %>%
  summarise(count = n()) 

crime_by_borough$population <- 
  with(crime_by_borough, ifelse(ARREST_BORO == "Brooklyn", population_boro[2,2],
                         ifelse(ARREST_BORO == "Bronx", population_boro[1,2],
                        ifelse(ARREST_BORO == "Queens", population_boro[4,2],
                         ifelse(ARREST_BORO == "Manhattan", population_boro[3,2],population_boro[5,2]      )))))

crime_by_borough <- crime_by_borough %>%
  group_by(ARREST_BORO) %>%
  mutate(offense_rate = (count/sum(count)))

# crime_by_borough <- crime_by_borough %>%
#   filter(crime_rate > 5000)
  
```

```{r}
#install.packages("gganimate")
#install.packages("gifski")

library(gganimate)
library(gifski)


p <- crime_by_borough %>%
  arrange(desc(offense_rate)) %>%
  group_by(ARREST_BORO) %>%
  slice(1:10) %>%
    ggplot(., aes(x = reorder(OFNS_DESC, offense_rate),
                                                 y = offense_rate,
                                                 fill = ARREST_BORO)) + 
    geom_bar(stat = "identity", show.legend = FALSE) +
    labs(title = "") +
    coord_flip()  + 
    xlab("Offense") +
    ylab("Offense Rate")

a <- p +
    transition_states(ARREST_BORO, wrap = FALSE)
animate(a, nframes = 5, fps = 1)
```

# Demographic of Perpetraitor in Each Borough

## Gender
```{r}
gender <- data_2018_2022 %>%
  group_by(PERP_SEX, ARREST_BORO)%>%
  summarise(count = n())

gender <- gender %>%
  group_by(ARREST_BORO) %>%
  mutate(percentage = (count/sum(count)))

ggplot(gender, aes (x = factor(percentage), y = percentage, fill = PERP_SEX))+
    geom_bar(stat="identity", na.rm = TRUE)+
    facet_wrap(~ARREST_BORO, ncol =3, scales = "free_x", drop = TRUE) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )
```
## Race
```{r}
race <- data_2018_2022 %>%
  group_by(PERP_RACE, ARREST_BORO)%>%
  summarise(count = n())

race <- race %>%
  group_by(ARREST_BORO) %>%
  mutate(percentage = (count/sum(count)))

ggplot(race, aes (x = factor(percentage), y = percentage, fill = PERP_RACE))+
    geom_bar(stat="identity", na.rm = TRUE)+
    facet_wrap(~ARREST_BORO, ncol =3, scales = "free_x", drop = TRUE) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )
```

# Source

1.  Interactive map <https://crimebythenumbers.com/choropleth-maps.html> <https://walker-data.com/census-r/mapping-census-data-with-r.html> <https://justinmorganwilliams.medium.com/basics-of-gis-mapping-with-r-using-grow-nyc-markets-75adcdd9b0> <https://cengel.github.io/R-spatial/mapping.html> <https://rpubs.com/schwarja209/NYC_hotspots>
